---
title: "R Notebook"
output: html_notebook
---

```{r}
#download.file("ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE71nnn/GSE71220/matrix/GSE71220_series_matrix.txt.gz","GSE71220_series_matrix.txt.gz")

# gene expression data: 617 people * 12381 genes
gene_expr = t(read.table("GSE71220_series_matrix.txt.gz",skip=66,nrows=12381)[,-1])
nsample = dim(gene_expr)[1]
n = dim(gene_expr)[2]

# which patients take statins
statin = read.table("GSE71220_series_matrix.txt.gz",skip=26,nrows=1)[-1]
statin = (strsplit(toString(unlist(statin)),"_")[[1]][2+3*(0:616)] == "statin")

# patients with COPD disease or healthy patients
disease = read.table("GSE71220_series_matrix.txt.gz",skip=37,nrows=1)[-1]
disease = (unlist(strsplit(strsplit(toString(unlist(disease)),": ")[[1]],","))[2*(1:nsample)] == "COPD")

# age of patient
age = read.table("GSE71220_series_matrix.txt.gz",skip=38,nrows=1)[-1]
age = as.numeric(unlist(strsplit(strsplit(toString(unlist(age)),": ")[[1]],","))[2*(1:nsample)])

# sex of patient (M or F)
sex = read.table("GSE71220_series_matrix.txt.gz",skip=39,nrows=1)[-1]
sex = (unlist(strsplit(strsplit(toString(unlist(sex)),": ")[[1]],","))[2*(1:nsample)])
```


```{r}
gene_list = which((cor(gene_expr[disease,],age[disease]))< -.2)
plot(cor(gene_expr[disease,gene_list],age[disease]),
     cor(gene_expr[disease,gene_list],statin[disease]),
     xlab='Corr with age', ylab='Corr with statin')
```



```{r}
dim(gene_expr[disease,gene_list])
length(age[disease])
length(cor(gene_expr[disease,gene_list],age[disease]))
```



```{r}
sample()
```









---
title: "STAT30850 MiniProject 1"
author: "Sarah Adilijiang, Hanyang Peng, Yanqing Gui"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.align='center', warning=FALSE, message=FALSE, 
                      echo=TRUE)
```

```{r}
library(dplyr)
```



## 1. Construct Data

```{r}
# load data
load("bikedata.RData")

# construct data frame
colnames(starttime) = c("year","month","day","hour","minute","second")
data = data.frame(station_start, station_end, duration, starttime,
                  day_of_week, days_since_Jan1_2010, member, bikenum)

# add a new column to identify a route
data["route"] = paste(data$station_start, seq="-", data$station_end)

dim(data) # 1342364 * 14
```

```{r}
# basic data summary
summary(data)  #bikenum has many NA's
```

```{r}
# aggregate data by route, day
new_data = data %>% group_by(station_start, station_end, year, month, day) %>% 
    summarise(average_duration=mean(duration), count=length(as.vector(duration)),
              route = unique(route), day_of_week=unique(day_of_week),
              days_since_Jan1_2010=unique(days_since_Jan1_2010))
nrow(new_data)  # 729321
```

After averaging the duration time for each day, there are 729321 data points.

From now on, we treat these daily average duration as our data points.

```{r}
# how many days have only one recorded duration
length(which(new_data$count==1)) # 432793
length(which(new_data$count==1)) / nrow(new_data) # 0.5934191
```

Among all the data, there are about 60% of the days have only one duration record for that day.




## 2. Select the potential routes with changes

### (1) Day count cutoff (keep: day count >5)

```{r}
# aggregate data by route
# and calculate the mean and standard deviation of average duration time
data_by_route = new_data %>% group_by(route) %>% 
    summarise(duration_mean=mean(average_duration), duration_sd=sd(average_duration),
              day_count=length(as.vector(average_duration)))
nrow(data_by_route)  # 12179
```

There are 12179 distinct routes in total.

```{r, fig.height=8, fig.width=10}
# check the day_counts of each route
summary(data_by_route$day_count)
par(mfrow=c(2,1))
hist(data_by_route$day_count, xlab="day_count", main="Histogram of day_counts of all routes",breaks=100)
hist(data_by_route$day_count[data_by_route$day_count<=100], xlab="day_count", main="Histogram of day_counts of routes (day_count<=100)", breaks=20)
```

The first high column is the routes with day_counts <= 5.

Because later we will permute the data within each group of {member, weekday}, which has four sub-groups for each route. To better compare, we only keep routes with day_counts over 4. And it also makes sense to remove the routes with low day_counts since it may have be occasional data which does not well represents the average duration change of one route.


```{r}
# keep the routes with day_counts >5
data_by_route = data_by_route[which(data_by_route$day_count>5), ]
nrow(data_by_route)  # 8187
```

After removing the routes with day_counts<=5, there are 8187 distinct routes left.




### (2) Duration time cutoff (keep: sd>300s=5min or sd/mean>1/3)

```{r, fig.height=8, fig.width=10}
# check the standard deviations of duration for each route
summary(data_by_route$duration_sd)

par(mfrow=c(2,1))
hist(data_by_route$duration_sd, xlab="sd_duration", main="Histogram of sd_duration of all routes", breaks=100)
hist(data_by_route$duration_sd[data_by_route$duration_sd<=2000], xlab="sd_duration", main="Histogram of sd_duration of routes (sd<=2000)", breaks=20)
```

Most of the routes have standard deviation of duration equal to or lower than 300 seconds, which is 5 minutes.

It makes sense to remove these routes since the standard deviation less than 5min indicates no significant time changes according to common sense for routes of general length. But for those short routes, 5min could be a large change. Therefore, we also keep the routes if the standard deviation is larger than 1/3 of the mean of duration.


```{r}
# keep the routes with (sd > 5min=300s) or (sd/mean > 1/3)
data_by_route = data_by_route[which(
    (data_by_route$duration_sd>300)|(data_by_route$duration_sd/data_by_route$duration_mean>1/3)), ]
nrow(data_by_route)  # 6641
12179 - 6641
```

After removing thee routes, there are finally 6641 distinct routes left, while 5538 routes have been removed from the orignal data.




### (3) Finally selected routes

```{r}
# all finally selected routes
routes = unique(data_by_route$route)

# keep only the selected routes in the origianl data
our_data = new_data[new_data$route %in% routes, ]
nrow(our_data)  # 655290
655290/729321
```


After removing the above routes from origal data, now there are 655290 data points left, which is about 90% of the original data.

Therefore, though we have removed many routes, most of the data points have been kept.






## 3. Permutation Test (not controlling for confounders)

```{r}
permutation_test = function(duration, day) {
    T_perms = rep(NA,500)  # 500 permutations
    for (k in 1:500) {
        duration_perm = sample(duration, replace=FALSE)
        T_perms[k] = abs(cor(duration_perm, day))
    }
    T_stat = abs(cor(duration, day))
    p_val = (1+sum(T_perms>=T_stat))/(1+500)
    return(p_val)
}
```


```{r}
set.seed(123)
results = matrix(NA,length(routes),4)
colnames(results) = c("route","station_start","station_end","p_value")
for (i in 1:length(routes)) {
    route_data = our_data[our_data$route==routes[i], ]
    results[i,1:3] = c(routes[i], route_data$station_start[1], route_data$station_end[1])
    results[i,4] = permutation_test(route_data$average_duration, route_data$days_since_Jan1_2010)
}
```

```{r}
# change the data type of the reported results
df_results = data.frame(results)
for (i in 1:3) {
    df_results[,i] = as.character(df_results[,i])
}
df_results$p_value = as.numeric(as.character(df_results$p_value))
str(df_results)
```

```{r}
# histogram of p-values
hist(df_results$p_value, xlab="p_value", main='p-values of permutation test', breaks=20)
```



```{r}
# routes with significant changes
sum(df_results$p_value<=0.05)  # 884 (level=0.05)
sum(df_results$p_value<=0.1)  # 1323 (level=0.1)

# select out the significant routes (level=0.05)  ?????????????????????????????
results_order = df_results[order(df_results$p_value), ]
results_sig = results_order[1:sum(df_results$p_value<=0.05), ]
```



```{r}
min_day = min(our_data$days_since_Jan1_2010)
max_day = max(our_data$days_since_Jan1_2010)

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = results_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(results_order)
for (i in 1:9) {
    route = results_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}
```


```{r}
# how many of signifance routes have same start station and end station
results_sig[results_sig$station_start==results_sig$station_end, ]
```






## 4. Permutation test (controlling for member)

```{r}
# aggregate data by route, day, member
member_data = our_data %>% group_by(station_start,station_end,year,month,day,member) %>% 
    summarise(average_duration=mean(duration), count=length(as.vector(duration)),
              route = unique(route), day_of_week=unique(day_of_week),
              days_since_Jan1_2010=unique(days_since_Jan1_2010))

# member_data = member_data[member_data$route %in% routes, ]
nrow(member_data)  # 726701
mean(member_data$member)  # 0.7953188
```

There are about 80% data points are the members' data.



```{r}
set.seed(123)
results_member = matrix(NA,length(routes),4)
colnames(results_member) = c("route","station_start","station_end","p_value")

for (i in 1:length(routes)) {
    route_data = member_data[member_data$route==routes[i], ]
    T_stat = abs(cor(route_data$average_duration, route_data$days_since_Jan1_2010))
    
    T_perms = rep(NA,500)  # 500 permutations
    for (k in 1:500) {
        if(sum(route_data$member)>1){
            route_data[route_data$member,"average_duration"] =
                sample(route_data[route_data$member,"average_duration"],replace=FALSE)
        }
        if(sum(!route_data$member)>1){
            route_data[!route_data$member,"average_duration"] = 
                sample(route_data[!route_data$member,"average_duration"],replace=FALSE)
        }
        T_perms[k] = abs(cor(route_data$average_duration, route_data$days_since_Jan1_2010))
    }
    results_member[i,4] = (1+sum(T_perms>=T_stat))/(1+500)
    results_member[i,1:3] = c(routes[i], route_data$station_start[1], 
                              route_data$station_end[1])
}
```


Reason of NA values: 

0 or 1 data point, or 2 data points with same duration time records for each sub-group.


```{r}
# change the data type of the reported results
df_results_member = data.frame(results_member)
for (i in 1:3) {
    df_results_member[,i] = as.character(df_results_member[,i])
}
df_results_member$p_value = as.numeric(as.character(df_results_member$p_value))
str(df_results_member)
```


```{r}
# histogram of p-values
hist(df_results_member$p_value, xlab="p_value", main='p-values of permutation test (controlling for member)', breaks=20)
```



```{r}
# routes with significant changes
sum(df_results_member$p_value<=0.05, na.rm=TRUE)  # 884 (level=0.05)
sum(df_results_member$p_value<=0.1, na.rm=TRUE)  # 1323 (level=0.1)

# select out the significant routes (level=0.05)  ?????????????????????????????
member_results_order = df_results_member[order(df_results_member$p_value), ]
member_results_sig = member_results_order[1:sum(df_results_member$p_value<=0.05, na.rm=TRUE), ]
```



```{r}
# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = member_results_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(results_order)
for (i in 1:9) {
    route = member_results_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}
```


```{r}
# how many of signifance routes have same start station and end station
member_results_sig[member_results_sig$station_start==member_results_sig$station_end, ]
```












## 5. Permutation test (controlling for weekday)

```{r}
# assign weekdays & weekends
our_data["weekday"] = TRUE
our_data[(our_data$day_of_week=="Sunday")|(our_data$day_of_week=="Saturday"), "weekday"] = FALSE

mean(our_data$weekday)  # 0.7953188
```

There are about ???? data points are the weekdays' data.



```{r}
set.seed(123)
results_weekday = matrix(NA,length(routes),4)
colnames(results_weekday) = c("route","station_start","station_end","p_value")

for (i in 1:length(routes)) {
    route_data = our_data[our_data$route==routes[i], ]
    T_stat = abs(cor(route_data$average_duration, route_data$days_since_Jan1_2010))
    
    T_perms = rep(NA,500)  # 500 permutations
    for (k in 1:500) {
        if(sum(route_data$weekday)>1){
            route_data[route_data$weekday,"average_duration"] =
                sample(route_data[route_data$weekday,"average_duration"],replace=FALSE)
        }
        if(sum(!route_data$weekday)>1){
            route_data[!route_data$weekday,"average_duration"] = 
                sample(route_data[!route_data$weekday,"average_duration"],replace=FALSE)
        }
        T_perms[k] = abs(cor(route_data$average_duration, route_data$days_since_Jan1_2010))
    }
    results_weekday[i,4] = (1+sum(T_perms>=T_stat))/(1+500)
    results_weekday[i,1:3] = c(routes[i], route_data$station_start[1], 
                               route_data$station_end[1])
}
```


Reason of NA values: 

0 or 1 data point, or 2 data points with same duration time records for each sub-group.


```{r}
# change the data type of the reported results
df_results_weekday = data.frame(results_weekday)
for (i in 1:3) {
    df_results_weekday[,i] = as.character(df_results_weekday[,i])
}
df_results_weekday$p_value = as.numeric(as.character(df_results_weekday$p_value))
str(df_results_weekday)
```


```{r}
# histogram of p-values
hist(df_results_weekday$p_value, xlab="p_value", main='p-values of permutation test (controlling for weekday)', breaks=20)
```



```{r}
# routes with significant changes
sum(df_results_weekday$p_value<=0.05, na.rm=TRUE)  # 884 (level=0.05)
sum(df_results_weekday$p_value<=0.1, na.rm=TRUE)  # 1323 (level=0.1)

# select out the significant routes (level=0.05)  ?????????????????????????????
weekday_results_order = df_results_weekday[order(df_results_weekday$p_value), ]
weekday_results_sig = weekday_results_order[1:sum(df_results_weekday$p_value<=0.05, na.rm=TRUE), ]
```



```{r}
# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = weekday_results_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(results_order)
for (i in 1:9) {
    route = weekday_results_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}
```


```{r}
# how many of signifance routes have same start station and end station
weekday_results_sig[weekday_results_sig$station_start==weekday_results_sig$station_end, ]
```










## 6. Permutation test (controlling for member and weekday)

```{r}
# assign weekdays & weekends
member_data["weekday"] = TRUE
member_data[(member_data$day_of_week=="Sunday")|(member_data$day_of_week=="Saturday"), "weekday"] = FALSE

mean(member_data$weekday)  # 0.7953188
```

There are about ???? data points are the weekdays' data.



```{r}
set.seed(123)
results_mb_wd = matrix(NA,length(routes),4)
colnames(results_mb_wd) = c("route","station_start","station_end","p_value")

for (i in 1:length(routes)) {
    route_data =member_data[member_data$route==routes[i], ]
    T_stat = abs(cor(route_data$average_duration, route_data$days_since_Jan1_2010))
    
    T_perms = rep(NA,500)  # 500 permutations
    member_weekday = (route_data$member) & (route_data$weekday)
    member_weekend = (route_data$member) & (!route_data$weekday)
    nonmember_weekday = (!route_data$member) & (route_data$weekday)
    nonmember_weekend = (!route_data$member) & (!route_data$weekday)
    for (k in 1:500) {
        if(sum(member_weekday)>1){
            route_data[member_weekday,"average_duration"] =
                sample(route_data[member_weekday,"average_duration"],replace=FALSE)
        }
        if(sum(member_weekend)>1){
            route_data[member_weekend,"average_duration"] =
                sample(route_data[member_weekend,"average_duration"],replace=FALSE)
        }
        if(sum(nonmember_weekday)>1){
            route_data[nonmember_weekday,"average_duration"] =
                sample(route_data[nonmember_weekday,"average_duration"],replace=FALSE)
        }
        if(sum(nonmember_weekend)>1){
            route_data[nonmember_weekend,"average_duration"] =
                sample(route_data[nonmember_weekend,"average_duration"],replace=FALSE)
        }
        T_perms[k] = abs(cor(route_data$average_duration, route_data$days_since_Jan1_2010))
    }
    results_mb_wd[i,4] = (1+sum(T_perms>=T_stat))/(1+500)
    results_mb_wd[i,1:3] = c(routes[i], route_data$station_start[1], 
                             route_data$station_end[1])
}
```


Reason of NA values: 

0 or 1 data point, or 2 data points with same duration time records for each sub-group.


```{r}
# change the data type of the reported results
df_results_mb_wd = data.frame(results_mb_wd)
for (i in 1:3) {
    df_results_mb_wd[,i] = as.character(df_results_mb_wd[,i])
}
df_results_mb_wd$p_value = as.numeric(as.character(df_results_mb_wd$p_value))
str(df_results_mb_wd)
```


```{r}
# histogram of p-values
hist(df_results_mb_wd$p_value, xlab="p_value", main='p-values of permutation test (controlling for member & weekday)', breaks=20)
```



```{r}
# routes with significant changes
sum(df_results_mb_wd$p_value<=0.05, na.rm=TRUE)  # 884 (level=0.05)
sum(df_results_mb_wd$p_value<=0.1, na.rm=TRUE)  # 1323 (level=0.1)

# select out the significant routes (level=0.05)  ?????????????????????????????
mb_wd_results_order = df_results_mb_wd[order(df_results_mb_wd$p_value), ]
mb_wd_results_sig = mb_wd_results_order[1:sum(df_results_mb_wd$p_value<=0.05, na.rm=TRUE), ]
```



```{r}
# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = mb_wd_results_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(results_order)
for (i in 1:9) {
    route = mb_wd_results_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration", type="l",
         xlim=c(min_day,max_day), main=route)
}
```


```{r}
# how many of signifance routes have same start station and end station
mb_wd_results_sig[mb_wd_results_sig$station_start==mb_wd_results_sig$station_end, ]
```











## 7. Combine results
```{r}
# double-check order of the routes in each result matrix
sum(df_results$route!=df_results_member$route)
sum(df_results$route!=df_results_weekday$route)
sum(df_results$route!=df_results_mb_wd$route)

# combine results
df_combine = cbind(df_results, df_results_member$p_value,
                   df_results_weekday$p_value, df_results_mb_wd$p_value)
colnames(df_combine)[4:7] = c("no_control","member","weekday","member_weekday")
str(df_combine)
```


```{r}
# report the number of significant cases (p<=0.05)
report = data.frame(group=c("no_control","member","weekday","member_weekday"),
                    sig_p=rep(NA,4),non_sig_p=rep(NA,4),NAs=rep(NA,4),total=rep(NA,4))
for (i in 1:4) {
    report[i,2] = sum(df_combine[,i+3]<=0.05, na.rm=TRUE)
    report[i,3] = sum(df_combine[,i+3]>0.05, na.rm=TRUE)
    report[i,4] = sum(is.na(df_combine[,i+3]))
    report[i,5] = sum(report[i,2:4])
}
report
```




```{r}
save('data','new_data','our_data','member_data', 'data_by_route', 'routes',
     'results','results_member','results_weekday','results_mb_wd',
     'df_results','df_results_member','df_results_weekday','df_results_weekend',
     'df_combine','report', 
     'min_day','max_day', file='v2_test.RData')
```



```{r}
save(data, new_data, our_data, member_data, data_by_route, routes,
     results, results_member, results_weekday, results_mb_wd,
     df_results, df_results_member, df_results_weekday, df_results_mb_wd,
     df_combine, report, 
     min_day, max_day, file='v2.RData')
```













































