---
title: "STAT30850 MiniProject 1"
author: "Sarah Adilijiang, Yanqing Gui, Hanyang Peng"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.align='center', warning=FALSE, message=FALSE, 
                      echo=TRUE, eval=FALSE)
```

save data

```{r}
save(data, new_data, our_data, member_data, routes, 
     T_stats, T_stats_member, results, df_results,
     df_lr, df_lr_member, df_lr_weekday, df_lr_member_weekday,
     df_lr_member2, df_lr_weekday2, df_lr_member_weekday2,
     file='MiniProject1.RData')
```

```{r,eval=TRUE}
load('MiniProject1.RData')
library(dplyr)
```







## 1. Construct Data

```{r}
# load data
load("bikedata.RData")

# construct data frame
colnames(starttime) = c("year","month","day","hour","minute","second")
data = data.frame(station_start, station_end, duration, starttime,
                  day_of_week, days_since_Jan1_2010, member, bikenum)

# add a new column to identify a route
data["route"] = paste(data$station_start, seq="-", data$station_end)

dim(data) # 1342364 * 14
```

```{r,eval=TRUE}
# basic data summary
summary(data)  #bikenum has many NA's
```

```{r}
# aggregate data by route, day
new_data = data %>% group_by(station_start, station_end, year, month, day) %>% 
    summarise(average_duration=mean(duration), count=length(as.vector(duration)),
              route = unique(route), day_of_week=unique(day_of_week),
              days_since_Jan1_2010=unique(days_since_Jan1_2010))
nrow(new_data)  # 729321
```

After averaging the duration time for each day, there are 729321 data points.

From now on, we treat these daily average duration as our data points.

```{r,eval=TRUE}
# how many days have only one recorded duration
length(which(new_data$count==1)) # 432793
length(which(new_data$count==1)) / nrow(new_data) # 0.5934191
```

Among all the data, there are about 60% of the days have only one duration record for that day.




## 2. Select the routes with potential changes

### (1) Day count cutoff (keep: day count >5)

```{r,eval=TRUE}
# aggregate data by route
# and calculate the mean and standard deviation of average duration time
data_by_route = new_data %>% group_by(route) %>% 
    summarise(duration_mean=mean(average_duration), duration_sd=sd(average_duration),
              day_count=length(as.vector(average_duration)))
nrow(data_by_route)  # 12179
```

There are 12179 distinct routes in total.

```{r, fig.height=8, fig.width=10, eval=TRUE}
# check the day_counts of each route
summary(data_by_route$day_count)
par(mfrow=c(2,1))
hist(data_by_route$day_count, xlab="day_count", main="Histogram of day_counts of all routes",breaks=100)
hist(data_by_route$day_count[data_by_route$day_count<=100], xlab="day_count", main="Histogram of day_counts of routes (day_count<=100)", breaks=20)
```

The first high column is the routes with day_counts <= 5.

Because later we will permute the data within each group of {member, weekday}, which has four sub-groups for each route. To better compare, we only keep routes with day_counts over 4. And it also makes sense to remove the routes with low day_counts since it may have be occasional data which does not well represents the average duration change of one route.


```{r, eval=TRUE}
# keep the routes with day_counts >5
data_by_route = data_by_route[which(data_by_route$day_count>5), ]
nrow(data_by_route)  # 8187
```

After removing the routes with day_counts<=5, there are 8187 distinct routes left.




### (2) Duration time cutoff (keep: sd>300s=5min or sd/mean>1/3)

```{r, fig.height=8, fig.width=10, eval=TRUE}
# check the standard deviations of duration for each route
summary(data_by_route$duration_sd)

par(mfrow=c(2,1))
hist(data_by_route$duration_sd, xlab="sd_duration", main="Histogram of sd_duration of all routes", breaks=100)
hist(data_by_route$duration_sd[data_by_route$duration_sd<=2000], xlab="sd_duration", main="Histogram of sd_duration of routes (sd<=2000)", breaks=20)
```

Most of the routes have standard deviation of duration equal to or lower than 300 seconds, which is 5 minutes.

It makes sense to remove these routes since the standard deviation less than 5min indicates no significant time changes according to common sense for routes of general length. But for those short routes, 5min could be a large change. Therefore, we also keep the routes if the standard deviation is larger than 1/3 of the mean of duration.


```{r, eval=TRUE}
# keep the routes with (sd > 5min=300s) or (sd/mean > 1/3)
data_by_route = data_by_route[which(
    (data_by_route$duration_sd>300)|(data_by_route$duration_sd/data_by_route$duration_mean>1/3)), ]
nrow(data_by_route)  # 6641
12179 - 6641
```

After removing thee routes, there are finally 6641 distinct routes left, while 5538 routes have been removed from the orignal data.




### (3) Finally selected routes

```{r}
# all finally selected routes
routes = unique(data_by_route$route)

# keep only the selected routes in the origianl data
our_data = new_data[new_data$route %in% routes, ]
nrow(our_data)  # 655290
655290/729321  # 0.8984933
```


After removing the above routes from origal data, now there are 655290 data points left, which is about 90% of the original data.

Therefore, though we have removed many routes, most of the data points have been kept.







## 3. Permutation Test (no controlling for confounders)

```{r}
# compute T_stat of all unique routes before permutation
# this is for easier use for permutation tests controlling for weekday
T_stats = data.frame(route=rep(NA,length(routes)), T_stat=rep(NA,length(routes)))
for (i in 1:length(routes)) {
    route_data = our_data[our_data$route==routes[i], ]
    T_stats[i,"route"] = routes[i]
    T_stats[i,"T_stat"] = abs(cor(route_data$average_duration, 
                                  route_data$days_since_Jan1_2010))
}
str(T_stats)
```


```{r}
set.seed(123)
results = matrix(NA,length(routes),4)
colnames(results) = c("route","station_start","station_end","no_control")

for (i in 1:length(routes)) {
    route_data = our_data[our_data$route==routes[i], ]
    results[i,1:3] = c(routes[i], route_data$station_start[1], route_data$station_end[1])
    
    duration = route_data$average_duration
    days = route_data$days_since_Jan1_2010
    
    T_perms = rep(0,500)  # 500 permutations
    for (k in 1:500) {
        duration_perm = sample(duration, replace=FALSE)
        T_perms[k] = abs(cor(duration_perm, days))
    }
    T = T_stats[T_stats$route==routes[i], "T_stat"]
    results[i,4] = (1+sum(T_perms>=T))/(1+500)
}
```

```{r}
# change the data type of the reported results
df_results = data.frame(results)
for (i in 1:3) {
    df_results[,i] = as.character(df_results[,i])
}
df_results$no_control = as.numeric(as.character(df_results$no_control))
str(df_results)
```

```{r,eval=TRUE}
# histogram of p-values
hist(df_results$no_control, xlab="p_value", 
     main='p-values of permutation test (no control)', breaks=20)
```



```{r,eval=TRUE}
# routes with significant changes
sum(df_results$no_control<=0.05)  # 893 (level=0.05)
sum(df_results$no_control<=0.1)  # 1324 (level=0.1)
```



```{r,eval=TRUE}
min_day = min(our_data$days_since_Jan1_2010)
max_day = max(our_data$days_since_Jan1_2010)

# order the p_values
results_order = df_results[order(df_results$no_control), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = results_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(results_order)
for (i in 1:9) {
    route = results_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```









## 4. Permutation test (controlling for weekday)

```{r}
# assign weekdays & weekends
our_data["weekday"] = TRUE
our_data[(our_data$day_of_week=="Sunday")|(our_data$day_of_week=="Saturday"), "weekday"] = FALSE

mean(our_data$weekday)  # 0.7428268
```

There are about 74% data points are the weekdays' data.


```{r}
df_results[,"weekday"] = rep(NA,nrow(df_results))

set.seed(123)
for (i in 1:nrow(df_results)) {
    route_data = our_data[our_data$route==df_results$route[i], ]
    
    duration = route_data$average_duration
    
    T_perms = rep(0,500)  # 500 permutations
    for (k in 1:500) {
        sim = route_data
        
        if(sum(route_data$weekday)>1){
            sim[sim$weekday,"average_duration"] = sample(duration[route_data$weekday])
        }
        if(sum(!route_data$weekday)>1){
            sim[!sim$weekday,"average_duration"] = sample(duration[!route_data$weekday])
        }
        
        T_perms[k] = abs(cor(sim$average_duration, route_data$days_since_Jan1_2010))
    }
    
    T = T_stats[T_stats$route==df_results$route[i], "T_stat"]
    df_results[i,"weekday"] = (1+sum(T_perms>=T))/(1+500)
    
    print(i)
}
```


```{r,eval=TRUE}
# histogram of p-values
hist(df_results$weekday, xlab="p_value", 
     main='p-values of permutation test (control for weekday)', breaks=20)
```



```{r,eval=TRUE}
# routes with significant changes
sum(df_results$weekday<=0.05, na.rm=TRUE)  # 830 (level=0.05)
sum(df_results$weekday<=0.1, na.rm=TRUE)  # 1238 (level=0.1)
```



```{r,eval=TRUE}
# order the p_values
weekday_order = df_results[order(df_results$weekday), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = weekday_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(weekday_order)
for (i in 1:9) {
    route = weekday_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```










## 5. Permutation test (controlling for member)

```{r}
# aggregate data by route, day, member
member_data = data %>% group_by(station_start,station_end,year,month,day,member) %>% 
    summarise(average_duration=mean(duration), count=length(as.vector(duration)),
              route = unique(route), day_of_week=unique(day_of_week),
              days_since_Jan1_2010=unique(days_since_Jan1_2010))
```

```{r}
# keep only the selected routes data
member_data = member_data[member_data$route %in% routes, ]
nrow(member_data)  # 726701
mean(member_data$member)  # 0.7953188
```

There are about 80% data points are the members' data.


```{r}
# compute T_stat of all unique routes before permutation (while seperating members & nonmembers)
# this is for easier use for permutation tests controlling for member & weekday
T_stats_member = data.frame(route=rep(NA,length(routes)), 
                            T_stat_member=rep(NA,length(routes)))
for (i in 1:length(routes)) {
    route_data = member_data[member_data$route==routes[i], ]
    T_stats_member[i,"route"] = routes[i]
    T_stats_member[i,"T_stat_member"] = abs(cor(route_data$average_duration, 
                                         route_data$days_since_Jan1_2010))
}
str(T_stats_member)
```


```{r}
df_results[,"member"] = rep(NA,nrow(df_results))

set.seed(123)
for (i in 1:nrow(df_results)) {
    route_data = member_data[member_data$route==df_results$route[i], ]
    
    duration = route_data$average_duration
    
    T_perms = rep(0,500)  # 500 permutations
    for (k in 1:500) {
        sim = route_data
        
        if(sum(route_data$member)>1){
            sim[sim$member,"average_duration"] = sample(duration[route_data$member])
        }
        if(sum(!route_data$member)>1){
            sim[!sim$member,"average_duration"] = sample(duration[!route_data$member])
        }
        
        T_perms[k] = abs(cor(sim$average_duration, route_data$days_since_Jan1_2010))
    }
    
    T = T_stats_member[T_stats_member$route==df_results$route[i], "T_stat_member"]
    df_results[i,"member"] = (1+sum(T_perms>=T))/(1+500)
    
    print(i)
}
```


```{r,eval=TRUE}
# histogram of p-values
hist(df_results$member, xlab="p_value", 
     main='p-values of permutation test (control for member)', breaks=20)
```



```{r,eval=TRUE}
# routes with significant changes
sum(df_results$member<=0.05, na.rm=TRUE)  # 726 (level=0.05)
sum(df_results$member<=0.1, na.rm=TRUE)  # 1146 (level=0.1)
```



```{r,eval=TRUE}
# order the p_values
member_order = df_results[order(df_results$member), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = member_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(member_order)
for (i in 1:9) {
    route = member_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```








## 6. Permutation test (controlling for member and weekday)

```{r}
# assign weekdays & weekends
member_data["weekday"] = TRUE
member_data[(member_data$day_of_week=="Sunday")|(member_data$day_of_week=="Saturday"), "weekday"] = FALSE

mean(member_data$weekday)  # 0.7299357
```

There are about 73% data points are the weekdays' data.


```{r}
df_results[,"member_weekday"] = rep(NA,nrow(df_results))

set.seed(123)
for (i in 1:nrow(df_results)) { 
    route_data = member_data[member_data$route==df_results$route[i], ]
    
    duration = route_data$average_duration
    
    T_perms = rep(0,500)  # 500 permutations
    for (k in 1:500) {
        sim = route_data
        
        if(sum((route_data$member) & (route_data$weekday))>1){
            sim[(sim$member) & (sim$weekday), "average_duration"] = sample(
            duration[(route_data$member) & (route_data$weekday)])
        }
        if(sum((route_data$member) & (!route_data$weekday))>1){
            sim[(sim$member) & (!sim$weekday), "average_duration"] = sample(
            duration[(route_data$member) & (!route_data$weekday)])
        }
        if(sum((!route_data$member) & (route_data$weekday))>1){
            sim[(!sim$member) & (sim$weekday), "average_duration"] = sample(
            duration[(!route_data$member) & (route_data$weekday)])
        }
        if(sum((!route_data$member) & (!route_data$weekday))>1){
            sim[(!sim$member) & (!sim$weekday), "average_duration"] = sample(
            duration[(!route_data$member) & (!route_data$weekday)])
        }
        
        T_perms[k] = abs(cor(sim$average_duration, route_data$days_since_Jan1_2010))
    }
    
    T = T_stats_member[T_stats_member$route==df_results$route[i], "T_stat_member"]
    df_results[i,"member_weekday"] = (1+sum(T_perms>=T))/(1+500)
    
    print(i)
}
```


```{r,eval=TRUE}
# histogram of p-values
hist(df_results$member_weekday, xlab="p_value", 
     main='p-values of permutation test (control for member & weekday)', breaks=20)
```



```{r,eval=TRUE}
# routes with significant changes
sum(df_results$member_weekday<=0.05, na.rm=TRUE)  # 692  (level=0.05)
sum(df_results$member_weekday<=0.1, na.rm=TRUE)  # 1130 (level=0.1)
```



```{r,eval=TRUE}
# order the p_values
member_weekday_order = df_results[order(df_results$member_weekday), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = member_weekday_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}


# plot several most non-significant routes
par(mfrow=c(3,3))
n = nrow(member_weekday_order)
for (i in 1:9) {
    route = member_weekday_order[n+1-i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```








## 7. Combine results

```{r,eval=TRUE}
# report the number of significant cases (p<=0.05)
report = data.frame(group=c("no_control","weekday","member","member_weekday"),
                    sig_p=rep(NA,4),non_sig_p=rep(NA,4),NAs=rep(NA,4),total=rep(NA,4))
for (i in 1:4) {
    report[i,2] = sum(df_results[,i+3]<=0.05, na.rm=TRUE)
    report[i,3] = sum(df_results[,i+3]>0.05, na.rm=TRUE)
    report[i,4] = sum(is.na(df_results[,i+3]))
    report[i,5] = sum(report[i,2:4])
}
report
```






## 8. Modified BH method to control FDR


```{r,eval=TRUE}
# modified BH method
modified_BH = function(P, alpha, gamma){
    n = length(P)
    order_index = order(P, decreasing=FALSE)
    pi_hat = sum(P>gamma)/n/(1-gamma)
    P[P>gamma] = Inf   # never reject any p > gamma
    ks = which(P[order_index] <= (1:n)*alpha/n/pi_hat)
    if(length(ks)==0){
        return(NULL)
    }else{
        k = max(ks)
        rej_index = order_index[1:k]
        return(rej_index)
    }
}
```


Like in Homework 2, we use a modified BH method to control FDR with level $\alpha=0.1$ and $\gamma=0.5$. So we aim to output a list of discoveries with FDP $\leq \alpha=0.1$.

```{r,eval=TRUE}
df_results[,"no_control_BH"] = rep(FALSE,nrow(df_results))
df_results[,"weekday_BH"] = rep(FALSE,nrow(df_results))
df_results[,"member_BH"] = rep(FALSE,nrow(df_results))
df_results[,"member_weekday_BH"] = rep(FALSE,nrow(df_results))

# FALSE = do NOT reject the null, 
# TRUE = reject the null, i.e. discovery with significant duration change
alpha = 0.1
gamma = 0.5
for (i in 1:4) {
    reject_index = modified_BH(df_results[,i+3], alpha, gamma)
    df_results[reject_index,i+7] = TRUE
}
```


```{r,eval=TRUE}
# report the number of significant cases (reject=TRUE in BH method)
report_BH = data.frame(group=c("no_control_BH","weekday_BH","member_BH","member_weekday_BH"), 
                       sig_p=rep(NA,4),non_sig_p=rep(NA,4),NAs=rep(NA,4),total=rep(NA,4))
for (i in 1:4) {
    report_BH[i,2] = sum(df_results[,i+7], na.rm=TRUE)
    report_BH[i,3] = sum(!df_results[,i+7], na.rm=TRUE)
    report_BH[i,4] = sum(is.na(df_results[,i+3]))
    report_BH[i,5] = sum(report_BH[i,2:4])
}
report_BH
```



```{r, eval=TRUE}
# significant routes determined by permutation test controlling for weekday
sig_perm_weekday = df_results[df_results$weekday_BH, ]
```








## 9. Linear Regression

### (1) Regress average duration only on date

```{r}
df_lr = data.frame(route=rep(NA,length(routes)), coef_date=rep(NA,length(routes)),
                   date_p=rep(NA,length(routes)))

for (i in 1:length(routes)) {
    route_data = our_data[our_data$route==routes[i], ]
    model = lm(average_duration~days_since_Jan1_2010, route_data)
    coefs = summary(model)$coefficients
    df_lr[i,1] = routes[i]
    df_lr[i,c("coef_date","date_p")] = c(coefs[2,1],coefs[2,4])
}
```


```{r}
# modified BH method for controlling FDR
df_lr[,"date_p_BH"] = FALSE
reject_index = modified_BH(df_lr[,"date_p"], alpha=0.1, gamma=0.5)
df_lr[reject_index,"date_p_BH"] = TRUE
```


```{r, eval=TRUE}
# histogram of p-values
hist(df_lr$date_p, xlab="p_value", 
     main='p-values of linear regression', breaks=20)
```

```{r, eval=TRUE}
# routes with significant changes
sum(df_lr$date_p<=0.05)  # 852  (level=0.05)
sum(df_lr$date_p_BH) # 180   (BH control)
```

```{r, eval=TRUE}
sig_lr = df_lr[df_lr$date_p_BH, ]
```


```{r,eval=TRUE}
# order the p_values
sig_lr_order = sig_lr[order(sig_lr$date_p), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = sig_lr_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```






### (2) Regress average duration on date, weekday

```{r}
df_lr_weekday = data.frame(route=rep(NA,length(routes)), coef_date=rep(NA,length(routes)),
                   date_p=rep(NA,length(routes)), coef_weeday=rep(NA,length(routes)),
                   weekday_p=rep(NA,length(routes)) )    

for (i in 1:length(routes)) { 
    route_data = our_data[our_data$route==routes[i], ]
    model = lm(average_duration~days_since_Jan1_2010 + weekday, route_data)
    coefs = summary(model)$coefficients
    df_lr_weekday[i,1] = routes[i]
    df_lr_weekday[i,c("coef_date","date_p")] = c(coefs[2,1],coefs[2,4])
    if(nrow(coefs)>2){
        df_lr_weekday[i,c("coef_weeday","weekday_p")] = c(coefs[3,1],coefs[3,4])
    }
}

index = which(is.na(df_lr_weekday$coef_weeday))
df_lr_weekday2 = df_lr_weekday[-index,]
```


```{r}
# modified BH method for controlling FDR
df_lr_weekday2[,"date_p_BH"] = FALSE
reject_index = modified_BH(df_lr_weekday2[,"date_p"], alpha=0.1, gamma=0.5)
df_lr_weekday2[reject_index,"date_p_BH"] = TRUE
```


```{r, eval=TRUE}
# histogram of p-values
hist(df_lr_weekday2$date_p, xlab="p_value", 
     main='p-values of linear regression (control for weekday)', breaks=20)
```

```{r, eval=TRUE}
# routes with significant changes
sum(df_lr_weekday2$date_p<=0.05)  # 807  (level=0.05)
sum(df_lr_weekday2$date_p_BH) # 120   (BH control)
```

```{r, eval=TRUE}
sig_weekday = df_lr_weekday2[df_lr_weekday2$date_p_BH, ]
sum(sig_weekday$weekday_p<=0.1)
sum(sig_weekday$weekday_p<=0.05)
```

```{r,eval=TRUE}
# order the p_values
sig_weekday_order = sig_weekday[order(sig_weekday$date_p), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = sig_weekday_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```






### (3) Regress average duration on date, member

```{r}
df_lr_member = data.frame(route=rep(NA,length(routes)), coef_date=rep(NA,length(routes)),
                   date_p=rep(NA,length(routes)), coef_member=rep(NA,length(routes)),
                   member_p=rep(NA,length(routes)) )    

for (i in 1:length(routes)) { 
    route_data = member_data[member_data$route==routes[i], ]
    model = lm(average_duration~days_since_Jan1_2010 + member, route_data)
    coefs = summary(model)$coefficients
    df_lr_member[i,1] = routes[i]
    df_lr_member[i,c("coef_date","date_p")] = c(coefs[2,1],coefs[2,4])
    if(nrow(coefs)==3){
        df_lr_member[i,c("coef_member","member_p")] = c(coefs[3,1],coefs[3,4])
    }
}

index = which(is.na(df_lr_member$coef_member))
df_lr_member2 = df_lr_member[-index, ]
```


```{r}
# modified BH method for controlling FDR
df_lr_member2[,"date_p_BH"] = FALSE
reject_index = modified_BH(df_lr_member2[,"date_p"], alpha=0.1, gamma=0.5)
df_lr_member2[reject_index,"date_p_BH"] = TRUE
```


```{r,eval=TRUE}
# histogram of p-values
hist(df_lr_member2$date_p, xlab="p_value", 
     main='p-values of linear regression (control for member)', breaks=20)
```

```{r, eval=TRUE}
# routes with significant changes
sum(df_lr_member2$date_p<=0.05)  # 633  (level=0.05)
sum(df_lr_member2$date_p_BH) # 77   (BH control)
```

```{r, eval=TRUE}
sig_member = df_lr_member2[df_lr_member2$date_p_BH, ]
sum(sig_member$member_p<=0.1)
sum(sig_member$member_p<=0.05)
```

```{r,eval=TRUE}
# order the p_values
sig_member_order = sig_member[order(sig_member$date_p), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = sig_member_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```







### (4) Regress average duration on date, member, weekday

```{r}
df_lr_member_weekday = data.frame(route=rep(NA,length(routes)), coef_date=rep(NA,length(routes)),
                   date_p=rep(NA,length(routes)), coef_member=rep(NA,length(routes)),
                   member_p=rep(NA,length(routes)), coef_weekday=rep(NA,length(routes)),
                   weekday_p=rep(NA,length(routes)) )    

for (i in 1:length(routes)) { 
    route_data = member_data[member_data$route==routes[i], ]
    model = lm(average_duration~days_since_Jan1_2010 + member + weekday, route_data)
    coefs = summary(model)$coefficients
    df_lr_member_weekday[i,1] = routes[i]
    df_lr_member_weekday[i,c("coef_date","date_p")] = c(coefs[2,1],coefs[2,4])
    if(nrow(coefs)==4){
        df_lr_member_weekday[i,c("coef_member","member_p")] = c(coefs[3,1],coefs[3,4])
        df_lr_member_weekday[i,c("coef_weekday","weekday_p")] = c(coefs[4,1],coefs[4,4])
    }
}

index = which(is.na(df_lr_member_weekday$coef_member))
df_lr_member_weekday2 = df_lr_member_weekday[-index, ]
```


```{r}
# modified BH method for controlling FDR
df_lr_member_weekday2[,"date_p_BH"] = FALSE
reject_index = modified_BH(df_lr_member_weekday2[,"date_p"], alpha=0.1, gamma=0.5)
df_lr_member_weekday2[reject_index,"date_p_BH"] = TRUE
```


```{r,eval=TRUE}
# histogram of p-values
hist(df_lr_member_weekday2$date_p, xlab="p_value", 
     main='p-values of linear regression (control for member & weekday)', breaks=20)
```

```{r, eval=TRUE}
# routes with significant changes
sum(df_lr_member_weekday2$date_p<=0.05)  # 637  (level=0.05)
sum(df_lr_member_weekday2$date_p_BH) # 62   (BH control)
```

```{r, eval=TRUE}
sig_member_weekday = df_lr_member_weekday2[df_lr_member_weekday2$date_p_BH, ]
sum((sig_member_weekday$member_p<=0.1) & (sig_member_weekday$weekday_p<=0.1))
sum((sig_member_weekday$member_p<=0.05) & (sig_member_weekday$weekday_p<=0.05))
```

```{r,eval=TRUE}
# order the p_values
sig_member_weekday_order = sig_member_weekday[order(sig_member_weekday$date_p), ]

# plot several most significant routes
par(mfrow=c(3,3))
for (i in 1:9) {
    route = sig_member_weekday_order[i,1]
    route_data = our_data[our_data$route==route, ]
    plot(route_data$days_since_Jan1_2010, route_data$average_duration, 
         xlab="days since Jan1 2020", ylab="average duration",
         xlim=c(min_day,max_day), main=route)
}
```







### (5) Combine results

```{r,eval=TRUE}
# report the number of significant cases (reject=TRUE in BH method)
report_lr = data.frame(group=c("no_control","weekday","member","member_weekday"), 
                       sig_p=rep(NA,4), sig_p_BH=rep(NA,4))
report_lr[1,2:3] = c(sum(df_lr$date_p<=0.05), sum(df_lr$date_p_BH))
report_lr[2,2:3] = c(sum(df_lr_weekday2$date_p<=0.05),sum(df_lr_weekday2$date_p_BH))
report_lr[3,2:3] = c(sum(df_lr_member2$date_p<=0.05),sum(df_lr_member2$date_p_BH))
report_lr[4,2:3] = c(sum(df_lr_member_weekday2$date_p<=0.05),sum(df_lr_member_weekday2$date_p_BH))
report_lr
```














